Lyman-$\alpha$ photons through rotating outflows

### 2018-08-16 paper/paper.tex
gaussian smoothing

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies. 
Here we introduce a semi-analytic model to quantify how rotating
outflows impact the morphology of the Lyman-$\alpha$ emission line.   
The model is contrasted against Monte Carlo radiative transfer
simulations of outflowing gas with additional solid body rotation.
We explore a range of neutral Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq
100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
For all viewing angles, the semi-analytic model reproduces the
radiative transfer results for the line width and flux change at the
line's center within a $7\%$ and $50\%$ precision for an optical depth of
$\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
depth of $\tauh=10^7$.
The semi-analytic model is a convenient tool to introduce rotational
kinematics as a post-processing step of idealized Monte Carlo
simulations, providing a framework to interpret \lya spectra in
systems where rotation is expected or directly measured through
kinematic maps.  
\end{abstract}

@@

\subsection{\lya continuum leakage or rotation?}

One application of this new model to the current interpretation of \lya
spectra, is that the central ($V=0$) emission of the spectra seen in the
\lya line, is a consequence of the viewing angle of the galaxy and can be
controlled by it. 
Several authors have suggested that this central emission is caused by
radiation that escapes the galaxy without scattering, mostly due to a
clumpy interstellar medium \citep{Hansen06, 2016ApJ...833L..26G}.
Rotation is an alternative that solves this issue.

We note that the relative intensity of the two peaks in the outflowing
spectra is not modified by rotation, that asymmetry is solely
controlled by the outflow velocity.
This means that results that have already derive a typical outflowing
velocity as to match the observational constraints do not have to be
completely revised. 
Only an additional exploration of rotational velocities and viewing
angles needs to be explored.

A recent observational example is provided by
\cite{2017A&A...608L...4R}.
They report optical spectroscopy of a bright lensed galaxy at a
redshift of $z=2.4$. 
The object presents a triple peaked \lya profile that can be thought
as the superposition of two components, a double asymmetric peak and a
narrower symmetric peak around the line's center.
One possible option for the narrow component is that it escaped the
galaxy without substantial interaction with neutral Hydrogen.
We speculate that the central feature could also be due to a purely 
rotating component that can produce such a feature as shown by
\cite{tololo}. The triple peak could be then produced by two main
decoupled components: one rotating the other outflowing.

Another interesting example are the \lya lines in Green Pea galaxies \citep{2016ApJ...820..130Y}.
\cite{2017ApJ...844..171Y} showed in a study of 43 Green Pea galaxies
that 2/3 were strong \lya emitters. 
There the \lya escape fraction is effectively estimated from the ratio
of H$\alpha$ to \lya.
Galaxies with large escape fractions also substantial \lya emission at
the line's center. 
We speculate that this feature could help to constrain rotation in
Green Pea galaxies and find new correlations between \lya escape
fraction and galaxy rotation.

@@

Although it is a good approximation to think the rotating spectra by a
sum of Doppler shifts, the peak of the spectra is also weighted by the
amount of mass with a given line-of-sight velocity. 
In this toy model the distance between the peaks of the
receding/approaching spectra is close to $\sim 20\ \kms$, which is
less than half of the naively expected value of $2\vrot=100\ \kms$,
due to the fact that only a small fraction of the photons are emitted
at the extreme of the galaxy having the maximum rotational velocity of
$50\ \kms$.

@@

\subsection{Low optical thickness regimes}
When located in the regimes of low optical thickness, \tauh $\lesssim 4$,  the agreement of the radiative transfer simulation with the semi-analytical solution is not as accurate. In this regime, only numerical simulations can help to predict the shape of the \lya profiles. We ran a simulation with low \tauh and compared it with the semi-analytical model. As seen in Figure \ref{fig:low_tau}, the model does not fit the simulation. This creates physical restrictions on the input of our model.

\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{low_tau_logtau4_theta90}
	\caption{\textbf{Low optical thickness regimes}.  
		Here we fix $\tauh=10^4$. 
		The six panels correspond to $\theta=90^\circ$.
		We vary \vrot increasing from left to right and \vout increasing
		from top to bottom. 
		The thin black line corresponds to the \lya line obtained with
		CLARA without any rotation and the indicated outflow velocity.
		The thick black line corresponds to the results including both
		outflows and rotation.
		The thick gray line shows the results of modifying the pure outflow
		solution by the Doppler shift presented in Equation \ref{eq:shift_x}
		(in thin line), if there is a radiative transfer of rotation and outflows
		(thick and clear line), and if there is a radiative transfer of
		only outflows, but also a Doppler shift from the rotational
		velocity (thick and dark line).  
		\label{fig:low_tau}}
\end{figure}

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies. 
Here we introduce a semi-analytic model to quantify how rotating
outflows impact the morphology of the Lyman-$\alpha$ emission line.   
The model is contrasted against Monte Carlo radiative transfer
simulations of outflowing gas with additional solid body rotation.
We explore a range of neutral Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq
100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
For all viewing angles, the semi-analytic model reproduces the
radiative transfer results for the line width and flux change at the
line's center within a $7\%$ and $50\%$ precision for an optical depth of
$\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
depth of $\tauh=10^7$.
The results of the semi-analytic model for $\tauh=10^6$ and
$\tauh=10^7$ can in turn be approximated by a gaussian smoothing with
standard deviation $\sigma_v = \vrot\sin(\theta)/2$. 
The semi-analytic model is a convenient tool to introduce rotational
kinematics as a post-processing step of idealized Monte Carlo
simulations, providing a framework to interpret \lya spectra in
systems where rotation is expected or directly measured through
kinematic maps.  
\end{abstract}

@@

\subsection{The semi-analytic model as gaussian smoothing}

The effects of the semi-analytic model on the pure outflowing spectra
are similar to the expected results from a gaussian smoothing.
this smoothing can also be a natural consequence of the intruments
used to measure the LAE spectra.
In other words, poor spectral resolution could be mistaken as an
indication for rotation.

We model the effects of gaussian smoothing with a similar ansatz as
the semi-analytic model by changing the frequency of each photon by a
Gaussian random variable centered on zero and width $\sigma_x$:
\begin{equation}
x^{\prime} = x + \mathcal{N}(0,\sigma_x).
\end{equation}

We find that the results of the semi-analytic model on the total
spectra at a given angle $\theta$ can be approximated using
$\sigma_x = (1/2)\times(\vrot/\vth)\times \sin(\theta)$. 
This approximation works well to reproduce the semi-analytic results
for $\tau=10^6$ and $\tau=10^7$.

@@

Although it is a good approximation to think the rotating spectra by a
sum of Doppler shifts, the peak of the spectra is also weighted by the
amount of mass with a given line-of-sight velocity. 
In this toy model the distance between the peaks of the
receding/approaching spectra is close to $\sim 25\ \kms$, which is
a fourth of the naively expected value of $2\vrot=100\ \kms$,
due to the fact that only a small fraction of the photons are emitted
at the extreme of the galaxy having the maximum rotational velocity of
$50\ \kms$.

@@

%\subsection{Low optical thickness regimes}
%When located in the regimes of low optical thickness, \tauh $\lesssim
%4$,  the agreement of the radiative transfer simulation with the
%semi-analytical solution is not as accurate. In this regime, only
%numerical simulations can help to predict the shape of the \lya
%profiles. We ran a simulation with low \tauh and compared it with the
%semi-analytical model. As seen in Figure \ref{fig:low_tau}, the model
%does not fit the simulation. This creates physical restrictions on the
%input of our model.

%\begin{figure}
%	\centering
%	\includegraphics[width=0.48\textwidth]{low_tau_logtau4_theta90}
%	\caption{\textbf{Low optical thickness regimes}.  
%		Here we fix $\tauh=10^4$. 
%		The six panels correspond to $\theta=90^\circ$.
%		We vary \vrot increasing from left to right and \vout increasing
%		from top to bottom. 
%		The thin black line corresponds to the \lya line obtained with
%		CLARA without any rotation and the indicated outflow velocity.
%		The thick black line corresponds to the results including both
%		outflows and rotation.
%		The thick gray line shows the results of modifying the pure outflow
%		solution by the Doppler shift presented in Equation \ref{eq:shift_x}
%		(in thin line), if there is a radiative transfer of rotation and outflows
%		(thick and clear line), and if there is a radiative transfer of
%		only outflows, but also a Doppler shift from the rotational
%		velocity (thick and dark line).  
%		\label{fig:low_tau}}
%\end{figure}
